### 2020-12-08 write/stress_rotation.tex

\documentclass[a4paper,12pt]{article}

\usepackage{amsmath} 
\usepackage{amssymb}

\title{Rotation of Reynolds stress tensor [for shear-alignment]} 
\begin{document} 
Let $u$, $v$ and $w$ (and the corresponding Reynolds stress tensor) 
be defined on a grid $Ox \times Oy \times Oz$ where the unit vector 
$\hat{e}_x$ is turned by $\theta$ with respect to the [negative] 
direction of surface shear stress \footnote{Here, negative means that we 
consider the grid aligned with the positive vector of the velocity gradient 
at the surface}
We require the velocities on the grid $Ox_\star \times Oy \times Oz_\star$ 
that is aligned with the surface shear stress, i.e. 
\begin{align}  
  \left(\begin{array}{c}
    \hat{e}_{x_\star}\\
    \hat{e}_y\\
    \hat{e}_{z_\star}
  \end{array} \right) = \left(\begin{array}{ccc}
    \cos\theta& 0 &-\sin\theta \\ 
    0 & 1 & 0 \\
    \sin\theta & 0 & \cos\theta 
  \end{array} \right)\left(\begin{array}{c} 
    \hat{e}_x \\ \hat{e}_y \\ \hat{e}_z
   \end{array}\right). 
\end{align} 
This implies 
\begin{eqnarray*} 
  u_\theta&=&u\cos\theta  +w\sin\theta \\ 
  w_\theta&=&-u\sin\theta +w\cos\theta. 
\end{eqnarray*}  
Let $c_\theta\equiv \cos\theta$, $s_\theta\equiv \sin_\theta$ and $m_\theta=\sin\theta\cos\theta$.
The corresponding Reynolds stress tensor is 
\begin{align} 
  \left(\begin{array}{ccc} 
    u_\theta u_\theta & u_\theta v_\theta & u_\theta w_\theta \\ 
    u_\theta v_\theta & v_\theta v_\theta & v_\theta w_\theta \\ 
    u_\theta w_\theta & v_\theta w_\theta & w_\theta w_\theta 
  \end{array} \right) =
  \left(\begin{array}{ccc} 
    (u c_\theta + w s_\theta )^2&
    uvc_\theta +wvs_\theta&
    (uc_\theta +ws_\theta)(-us_\theta +wc_\theta)\\
    & vv & (-us_\theta+wc_\theta)v\\ 
    && (-us_\theta +w c_\theta)^2 \end{array}\right)\\
  = \left(\begin{array}{ccc}  
    u^2c_\theta^2 +w^2 s_\theta^2 + 2uw m_\theta&
    uvc_\theta +wvs_\theta&
    m_\theta (w^2-u^2)+uw(c_\theta^2 -s_\theta^2) \\
    &v^2& -uvs_\theta + wvc_\theta\\
    && u^2s_\theta^2 + w^2 c_\theta^2 - 2uw m_\theta
    \end{array}\right) \\ 
  = \left(\begin{array}{ccc} 
    u^2c_\theta +w^2s_\theta
  \end{array}
\end{align} 
\end{document}